\chapter{Introduction}

This file contains the basic formatting instructions and guidelines on how to write your dissertation, according to the School's requirements. Word Styles are being used extensively for the formatting of basic elements, such as titles, legends etc. Minor modifications are permitted with the agreement of the supervisor. \par
This file is also a sample of how your dissertation should look like in the end. In order to avoid distortions and alterations, it is strongly suggested that when you write, you should have control character ¶ always on, by clicking on the button shown below on your toolbar:

\section{General guidelines - pagination}
The final text will be printed on both sides of A4 size paper. All page margins (top, bot-tom, left, right) should be set at 25mm. Additionally, you should allow 10mm for gutter. The text should be fully justified and you should use hyphenation in order to eliminate unnecessary big gaps between words (these settings have already been set in this document). \par
For Normal text use Times New Roman font, size 11 or 12 pts and line spacing of 1½ line. To emphasize a word you should use italics, not bold or underlined. 

\subsection{Pictures}
Pictures are added in-line, with a caption below them, near the paragraph text where they are referenced (usually, shortly after). All pictures should be referenced at least once in the paragraph text. References could be made using the right MS Word shortcuts (cross-reference). This is an example of this: Picture 1 shows two examples of palmtops.

\subsection{Pictures}
Table captions are added above the table. After the table leave a blank line, as in the example. \par
Same as with pictures, tables should be referenced at least once in the paragraph text. Cross-referencing is encouraged here too.
